---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
First, unzip the provided zip file in the current directoy, and then read it.
```{r }
unzip("activity.zip", overwrite = TRUE)
rawData <- read.csv("activity.csv")
```

Second, preprocess raw data.
Steps column is as-is.
Convert string representation of date to Date class.
Convert integer representation of time of the day to minutes since midnight.
For example, integer value 1155 in raw data means 11:55 of the day.
```{r }
preprocessedData <- data.frame(steps = rawData$steps)
preprocessedData$date <- as.Date(rawData$date, "%Y-%m-%d")
preprocessedData$interval <- rawData$interval %/% 100 * 60 + rawData$interval %% 100
```

Display a few observations.
```{r }
head(preprocessedData)
tail(preprocessedData)
```

Display the summary.
```{r }
summary(preprocessedData)
```

## What is mean total number of steps taken per day?
Calculate the total number of steps taken each day.
```{r }
totalStepsPerDay <- tapply(preprocessedData$steps, preprocessedData$date, sum)
```

Draw histogram.
```{r }
hist(totalStepsPerDay)
```

Calucate the mean total number of steps taken per day.
```{r }
mean(totalStepsPerDay, na.rm = TRUE)
```

Calculate the median total number of steps taken per day.
```{r }
median(totalStepsPerDay, na.rm = TRUE)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
